---
title: "Analysis"
output: html_document
---
```{r}
library("deSolve", lib.loc="/usr/local/lib/R/3.5/site-library")
```

```{r}
A_=1
bcr_=10
bcrini_=4
bcrend_=4.1
cd40_=5
cd40ini_=5
cd40end_=5.125
typeGRN_="CMYCFOXO"
xmax_=20
ymax_=10
xsteps_=0.01
typePars_="CMYCFOXO"
```

#In setPars function, define parameters as function of the type of model desired to run (type_) which can be the healthy or any of the lymphoma models describedin Martinez et al., 2012. 
#Affinity and bcr/cd40 signals can also be defined externally (A_,bcr_,cd40_) to be able to call this function at several timesteps and change theirvalue acording to protocol (see experiment function).
#Note: When modeling any of the lymphoma model parameters below always use setGRN type= "GRN".
```{r}
setPars<-function(A_,bcr_,cd40_, type_){
  
  if(type_ == "Healthy") # simplest Martinez model IRF4-->IRF4
  {
    
    pars<-c(MuP=1E-6, MuB=2, MuI=0.1,
            SigmaP=9, SigmaB=100, SigmaI=2.6, 
            kP=1, kB=1, kI=1, 
            LambdaP=1, LambdaB=1, LambdaI=1, 
            bcr0=bcr_ ,cd400=cd40_, A=A_)
    return(pars) 
  }
    else if(type_ == "M8")
  {
    pars<-c(MuP=1E-6, MuB=2
            , MuI=0.1*2 #M8 model
            ,SigmaP=9, SigmaB=100, SigmaI=2.6*2, #M8 model
            kP=1, kB=1, kI=1, 
            LambdaP=1, LambdaB=1, LambdaI=1, 
            bcr0=bcr_ ,cd400=cd40_, A=A_)
    return(pars) 
  }
   else if(type_ == "CMYCFOXO")
  {
    pars<-c(MuP=1E-6, MuB=2, MuI=0.1,MuF=1,MuC=1E-6,
            SigmaP=9, SigmaB=100, SigmaI=2.6,SigmaF=2.6,SigmaC=2.6,
            kP=1, kB=1, kI=1,kF=1,kC=1,  
            LambdaP=1, LambdaB=1, LambdaI=1, LambdaF=1,LambdaC=1,
            bcr0=bcr_ ,cd400=cd40_, A=A_)
    return(pars) 
   }
     else if(type_ == "CMYCFOXO_M8")
  {
    pars<-c(MuP=1E-6, MuB=2, MuI=0.1*2 #M8 model
            ,MuF=1,MuC=1E-6,
            SigmaP=9, SigmaB=100, SigmaI=2.6*2 #M8 model
            ,SigmaF=2.6,SigmaC=2.6,
            kP=1, kB=1, kI=1,kF=1,kC=1,  
            LambdaP=1, LambdaB=1, LambdaI=1, LambdaF=1,LambdaC=1,
            bcr0=bcr_ ,cd400=cd40_, A=A_)
    return(pars) 
  }
}
```

```{r}
yini<-function(type_){
  if(type_ == "CMYCFOXO"||type_ == "CMYCFOXO_M8")
  {
    yini<-c(IRF4=0,CMYC=0,FOXO1=1,BLIMP1=0,BCL6=5)  # ORDER OF INITIAL CONDITIONS SHOULD ALWAYS MATCH ORDER OF ODE EQUATIONS!
  }
  else
  {
    yini<-c(BLIMP1=0,BCL6=5,IRF4=0) 
  }
  return(yini)
}
```



#In setGRN function, define desired model to run (type_) which can be the healthy or any of the lymphoma models describedin Martinez et al., 2012. 
#Note: When modeling any of the lymphoma models below always use "Healthy" parameters.
```{r}
setGRN<-function(type_)
{
  if(type_ == "GRN") # simplest Martinez model
  {
    GRN<-function(t,y,pars){
      with(as.list(c(y,pars)), {
        dBLIMP1 <- MuP + SigmaP*kB^2/(kB^2 + BCL6^2) + SigmaP*IRF4^2/(kI^2 + IRF4^2) - LambdaP*BLIMP1 
        dBCL6 <- MuB + SigmaB*kB^2/(kB^2 + BCL6^2)*kP^2/(kP^2 + BLIMP1^2)*kI^2/(kI^2 + IRF4^2) - (LambdaB + bcr0*kB^2/(kB^2 + BCL6^2))*BCL6
        dIRF4 <- MuI + SigmaI*IRF4^2/(kI^2 + IRF4^2) + (cd400*A)*kB^2/(kB^2 + BCL6^2) - LambdaI*IRF4
        return(list(c(dBLIMP1,dBCL6,dIRF4)))
      })}
  }
  else if(type_ == "CMYCFOXO") # simplest Martinez model
  {
    GRN<-function(t,y,pars){
      with(as.list(c(y,pars)), {
        dIRF4 <- MuI + SigmaI*IRF4^2/(kI^2 + IRF4^2)  + (cd400*A)*kB^2/(kB^2 + BCL6^2) - LambdaI*IRF4
        dCMYC <- MuC + SigmaC*(kF^2/(kF^2+FOXO1^2))  + SigmaC*(kP^2/(kP^2+BLIMP1^2)) + SigmaC*IRF4^2/(kI^2 + IRF4^2) - LambdaC*CMYC
        dFOXO1 <- MuF - bcr0  -LambdaF*FOXO1 -(cd400*A)
        dBLIMP1 <- MuP + SigmaP*kB^2/(kB^2 + BCL6^2) + SigmaP*IRF4^2/(kI^2 + IRF4^2) + SigmaP*FOXO1^2/(kF^2 + FOXO1^2) - LambdaP*BLIMP1 
        dBCL6 <- MuB + SigmaB*kB^2/(kB^2 + BCL6^2)*kP^2/(kP^2 + BLIMP1^2)*kI^2/(kI^2 + IRF4^2) - (LambdaB + bcr0*kB^2/(kB^2 + BCL6^2))*BCL6
        return(list(c(dIRF4,dCMYC,dFOXO1,dBLIMP1,dBCL6)))
      })}
  }
  return(GRN)
  }
```

# In Experiment function, with given time boundaries of simulation as well as affinity and bcr/cd40 maximal values define affinity and bcr/cd40 values at each time step. 
# Also define initial values (yini) and use ode to solve the system of equations at each time step.
# Return a data frame with the calculated levels of BLIMP1,BCL6,IRF4,A,bcr,cd40. 
```{r}
experiment<-function(A_,bcr_,bcrini_,bcrend_,cd40_,cd40ini_,cd40end_,typeGRN_,xmax_,xsteps_,typePars_)
{
    times<-seq(from=0, to=xmax_-xsteps_, by= xsteps_)
    
    A<- c(rep(0, (xmax_/xsteps_)))# vector of zeros 
    A <- replace(A,list((bcrini_/xsteps_):(bcrend_/xsteps_))[[1]],A_)#insert signals =1
    A <- replace(A,list((cd40ini_/xsteps_):(cd40end_/xsteps_))[[1]],A_)#insert signals =1
    
    bcr <- c(rep(0, (xmax_/xsteps_)))
    bcr <- replace(bcr,list((bcrini_/xsteps_):(bcrend_/xsteps_))[[1]],bcr_)
    
    cd40 <- c(rep(0, (xmax_/xsteps_)))
    cd40 <- replace(cd40,list((cd40ini_/xsteps_):(cd40end_/xsteps_))[[1]], cd40_)
    
    #Initial condition at time=0
    yini<- yini(typePars_)
    
    #Use first two elements of times as first point (times[1])
    data<-data.frame(ode(yini,c(times[[1]],times[[2]]), setGRN(typeGRN_), setPars(A[[2]],bcr[[2]],cd40[[2]],typePars_)))
     #Basal BCR and CD40 signal
    for (i in seq_along(times))
    {
      if(i > 2)
      {
        #replace yini (TF levels for next step) from previous timestep
        yini<-c(replace(yini,list(seq_along(data[-1]))[[1]],as.numeric(data[i-1,][-1])))
        
        # Use  TF levels of previous time point as yini (ode needs at least 2 time points: use 0 and i only take i)
        ODE<-data.frame(lsode(yini,c(times[[i-1]],times[[i]]), setGRN(typeGRN_), setPars(A[[i]],bcr[[i]],cd40[[i]],typePars_)))
        data[i,]<-ODE[-1,]#overright data with second time point which will then bbe used as yini in next iteration
        A[[i]]<-A[[i]]
      }
    }
    data$Affinity<-A
    data$BCR<-bcr
    data$CD40<-cd40
  return(data)
}
```

#Plot results returned from experiment(Colour).
```{r}
plotResults<-function(A_,bcr_,bcrini_,bcrend_,cd40_,cd40ini_,cd40end_,typeGRN_,xmax_,xsteps_,ymax_,typePars_)
{
  data_<-experiment(A_,bcr_,bcrini_,bcrend_,cd40_,cd40ini_,cd40end_,typeGRN_,xmax_,xsteps_,typePars_)
  pars<- setPars(A_,bcr_,cd40_,typePars_)
  #list of colours for lines and legend
  colourlist<-c("green","cyan","blue","black","red","orange","blue","gold")
  plot(data_$time, data_[[2]],xlab="time (4hrs)",
       ylab="Concentration (10^-8 M)",
       type="l", col=colourlist[[1]], 
       ylim=c(0,max(ymax_)),xlim=c(0,max(xmax_)))
  # Skip times and first transcription factor in data to plot remaining lines
  for (i in 2:(length(data_)-2))
  {
    lines(data_$time,data_[[i]],col=colourlist[[i-1]])
  }
  lines(data_$time,data_[[length(data_)-1]],col=colourlist[[6]])  
  lines(data_$time,data_[[length(data_)]],col=colourlist[[7]])
  legend(xmax_-2,ymax_,c(names(data_)[2],names(data_)[3],names(data_)[4],names(data_)[5],names(data_)[6]), text.col = c(colourlist[[1]],colourlist[[2]],colourlist[[3]],colourlist[[4]],colourlist[[5]]),bty="n")
  #  axis(4,at= c(0,cd40_),labels=c(0,cd40_),las=0)
 # mtext("Signaling ", side=4, line=-1.5, col="black" ,las=0)

}
```

#Plot results returned from experiment (Black and White).
```{r}
plotResults_bw<-function(A_,bcr_,bcrini_,bcrend_,cd40_,cd40ini_,cd40end_,typeGRN_,xmax_,xsteps_,ymax_,typePars_)
{
  data_<-experiment(A_,bcr_,bcrini_,bcrend_,cd40_,cd40ini_,cd40end_,typeGRN_,xmax_,xsteps_,typePars_)
  pars<- setPars(A_,bcr_,cd40_,typePars_)
  #list of types of plot drawn for lines and legend
  ltylist<-c(2,3,4,5,1,1,1)
  plot(data_$time, data_[[2]], xlab="time (4hrs)",
       ylab="Concentration (10^-8 M)",
       lty =ltylist[[1]], col="black", type="n",
       ylim=c(0,max(ymax_)),xlim=c(0,max(xmax_)),bty="n")
  # Skip times and first transcription factor in data to plot remaining lines
  for (i in 2:(length(data_)-2))
  {
    lines(data_$time,data_[[i]],lty=ltylist[[i-1]],col="black")
  }
  lines(data_$time,data_[[length(data_)-1]]/5,lty=ltylist[[6]],col="black")
  lines(data_$time,data_[[length(data_)]]/5,lty=ltylist[[6]],col="black")
  legend(xmax_-5,ymax_,c(names(data_)[2],names(data_)[3],names(data_)[4],names(data_)[5]), lty=  c(ltylist[[1]],ltylist[[2]],ltylist[[3]],ltylist[[4]],ltylist[[5]]) ,text.col = "black" , bty="n")
 # axis(4,at= c(0,cd40_),labels=c(0,cd40_),las=0)
 # mtext("Signaling ", side=4, line=-1.5, col="black" ,las=0)
}
```


```{r}
#plotting Antoines Network (with his parameters)
jpeg("AntoinesBCLnetwork.jpeg", width = 5, height = 4, units = 'in', res = 300)
plotResults(0,12,45,45.1,5,60,60.1,"CMYCFOXO",200,0.1,15,"CMYCFOXO")
dev.off()

```


```{r}
jpeg("Healthy_A1.jpeg", width = 5, height = 4, units = 'in', res = 300)
plotResults(1,1,4,4.1,50,5,5.125,"GRN",15,0.01,15,"Healthy")
dev.off()
jpeg("M8_A1.jpeg", width = 5, height = 4, units = 'in', res = 300)
plotResults(1,1,4,4.1,50,5,5.125,"GRN",15,0.01,15,"M8")
dev.off()
```
```{r}
#CC recycling takes 10 hrs (MYC increase)
#PC differentiation can take between 10 and 34 hours (recycling + divisions)

jpeg("CMYCFOXO_A0.jpeg", width = 5, height = 4, units = 'in', res = 300)
plotResults(0,1,4,4.1,50,5,5.125,"CMYCFOXO",14,0.1,15,"CMYCFOXO")
dev.off()
jpeg("CMYCFOXO_A0.5.jpeg", width = 5, height = 4, units = 'in', res = 300)
plotResults(0.5,1,4,4.1,50,5,5.125,"CMYCFOXO",8,0.1,15,"CMYCFOXO")
plotResults(0.5,1,4,4.1,50,5,5.125,"CMYCFOXO",14,0.1,15,"CMYCFOXO")
dev.off()
jpeg("CMYCFOXO_A1.jpeg", width = 5, height = 4, units = 'in', res = 300)
plotResults(1,1,4,4.1,50,5,5.125,"CMYCFOXO",8,0.1,15,"CMYCFOXO")
dev.off()

```


```{r}
jpeg("CMYCFOXO_M8_A0.jpeg", width = 5, height = 4, units = 'in', res = 300)
plotResults(0,1,4,4.1,50,5,5.125,"CMYCFOXO",8,0.1,15,"CMYCFOXO_M8")
dev.off()
jpeg("CMYCFOXO_M8_A0.5.jpeg", width = 5, height = 4, units = 'in', res = 300)
plotResults(0.5,1,4,4.1,50,5,5.125,"CMYCFOXO",8,0.1,15,"CMYCFOXO_M8")
dev.off()
jpeg("CMYCFOXO_M8_A1.jpeg", width = 5, height = 4, units = 'in', res = 300)
plotResults(1,1,4,4.1,50,5,5.125,"CMYCFOXO",8,0.1,15,"CMYCFOXO_M8")
dev.off()
```

